---
title: "HW04 Part 1: Complete Chapter 3"
author: "Melvin Cummings"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
library(tidyverse)
```

#### Plot <no.1>
Plot type: scatterplot
Dataset: trees
x-axis: height
y-axis: girth
```{r}
data(trees)
?trees
ggplot(data = trees) + geom_point(mapping = aes(x=Height, y = Girth))
```

This plot shows a positive correlation between the girth and height of trees.
#### Plot <no.2>

```{r}
?year
year <- c(1821:1934)

class(lynx)
as.vector(year)

pelts <- c(as.vector(lynx))
pelts

lynx_pelts <- data.frame(year, pelts)
lynx_pelts

ggplot(data = lynx_pelts, mapping = aes(x = year, y = pelts)) + geom_line(color = "maroon")
```
This plot shows the treadline and difference in years as pelts increases and decreases.

#### Plot <no.3>
Plot type: scatterplot
Dataset: iris
x-axis: petal length
y-axis: petal width
Point color and shape by species. You do not have to use fillable shapes.
Point size of 2
Add a labs layer to change the x- and y-axis labels so that they do not have periods in the names (i.e., Petal Length, Petal Width).

```{r}
data("iris")
?iris
ggplot(data = iris) + geom_point(mapping = aes(x = Petal.Length, y = Petal.Width, color = Species, shape = Species, size = 2)) + labs(x = "Petal Length") + labs(y = "Petal Width")


```

This plot shows the correlation between the petal length and petal width of the differenct species of iris
#### Plot <no.4>
 Plot type: Violin plot (look up geom_violin)
• Dataset: msleep
• x-axis: vore
• y-axis: sleep_rem
• fill = grayXX where XX is either 30 or 70.
• In your description, describe in your own words what violin plots display (you can search the interwebs),
and what is the difference among the two versions of gray shading. Hint: the grays extend from gray0
to gray100. You can learn more about colors in R from
```{r}
data("msleep")
?msleep
ggplot(msleep) + geom_violin(mapping = aes(x = vore, y = sleep_rem), fill = "grey30")
```

This plot shows the correlation between the rem sleep and what type of mammal.
#### Plot <no.5>
 Plot type: Violin plot (look up geom_violin)
• Dataset: msleep
• x-axis: vore
• y-axis: sleep_rem
• fill = grayXX where XX is either 30 or 70.
• In your description, describe in your own words what violin plots display (you can search the interwebs),
and what is the difference among the two versions of gray shading. Hint: the grays extend from gray0
to gray100. You can learn more about colors in R from
```{r}
data("msleep")
?msleep
ggplot(msleep) + geom_violin(mapping = aes(x = vore, y = sleep_rem), fill = "grey70")
```
This plot is nearly identical to plot 4 but has a different shade of gray when compared to plot 4.
#### Plot <no.6>
#### Plot <no.7>
• Plot type: boxplot
• Dataset: msleep
• x-axis: order
• y-axis: sleep_total
• use coord_flip()

```{r}
data("msleep")
?msleep
ggplot(data = msleep) + geom_boxplot(mapping = aes(x = order, y = sleep_total)) + coord_flip()

```

This plot is a box plot which shows the order of mammals when compared to total sleep
#### Plot <no.8>
Plot type: boxplot with points
• Dataset: msleep
• x-axis: conservation
• y-axis: awake
3
• You must have a boxplot layer, a point layer, and a jitter layer.
• color by conservation status.
• You may use coord_flip but it is not required. Try both and choose the one you think looks best.
• Add a lab layer to change both axis labels so each starts with an upper-case letter.
• Search the internet to find out how to change the legend title to Conservation. Make that change.
(Do not try to change the actual legend entries like “cd” and “vu”). Note: This can be done a couple
of different ways but scale_color_discrete() is one good way

```{r}
data("msleep")
ggplot(data = msleep) + geom_boxplot(mapping = aes(x = conservation, y = awake, color = conservation)) + labs(x = "Conservation") + labs(y = "Awake") + geom_jitter(mapping = aes(x = conservation, y = awake, color = conservation)) + geom_point(mapping = aes(x = conservation, y = awake, color = conservation)) + coord_flip() + scale_fill_discrete(name = "Conservation")

```

This plot shows the time that the mammal is awake when compared to conservation.
#### Plot <no.9>
One should plot any one of the “sleep” or “awake” variables against body weight. The other should
plot any one of the “sleep” or “awake” variables against brain weight.
• In each plot, color the points by one of the nominal data categories. Use a different category for each
plot.
• Apply facet_wrap() to at least one of the plots using one of the nominal variables. You decide whether
you use 2 or 3 columns. Hint: use one of the nominal variables with relatively few different types for
wrapping. Explore: What happens if you use a nominal variable like genus, with lots of types?
• You should try a few versions of each graph until you find combinations that you think show some clear
trends.

• Describe the patterns or trends you see in each graph.
If you use nominal the data ruins the graph by being to large
```{r}
ggplot(data = msleep) + geom_point(mapping = aes(x = awake, y = brainwt, color = sleep_total)) 
```

This plot shows the amount of mammal awake when compared to the brain weight.
#### Plot <no.10>

```{r}
ggplot(data = msleep) + geom_point(mapping = aes(x = sleep_rem, y = bodywt, color = sleep_cycle)) + facet_wrap(~ vore, ncol = 2) 
```

These plots shows the difference of rem sleep when compared to body weight between the different mammal vore types.