---
title: "HW04 Part 1: Complete Chapter 3"
author: "Melvin Cummings"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Chapter 3

- Change "your name" in the YAML header above to your name.

### Section 3.1

Load the tidyverse packages. I've entered the first code chunk for you. **You must enter all subsequent code chunks and run the code.** 

```{r}
library("tidyverse")
```

If you get an error, then you did not install the `tidyverse` package like you were supposed to in the previous assignment. Go back to that assignment and complete that step. (This should not be the case, though.)

### Section 3.2: First steps

Enter your code chunks for Section 3.2 here.

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy))

```



Get in the habit of entering a brief description of what each chunk of code does. Enter this above the chunk, like I did in Section 3.1.

#### Section 3.2 Questions

Answer the questions *completely.* Some answers will require you to write code in a code chunk and also to type text before or after the chunk. Other answers will require only code or only text. 

**1:** Run ggplot(data = mpg). What do you see?
An empty graph

```{r}
ggplot(data = mpg)
```




**2:** How many rows are in `mpg`? How many columns? **Hint:** Use the `dim()` function. Type `dim()` in the console to learn what `dim()` does.

Rows: 234   Columns:11

```{r}
dim(mpg)
```


**3:** What does the `drv` variable describe? Read the help for `?mpg` to find out.

f = front-wheel drive, r = rear wheel drive, 4 = 4wd

**4:** Make a scatterplot of `hwy` vs `cyl`.

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = hwy, y = cyl))
```


**5:** What happens if you make a scatterplot of `class` vs `drv`? Why is the plot not useful?

This plot isnt useful because one class of car can have multiple wheel drive, such as subcompact having front,  rear, and 4 wheel drive

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = class, y = drv))
```



### Section 3.3: Aesthetic mappings


#### Section 3.3 questions

**1:** What’s gone wrong with this code? Why are the points not blue?

The points are not blue since colour is found within the aes, mapping and is consided as one value

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, colour = "blue"))
```


**2:** Which variables in `mpg` are categorical? Which variables are continuous? (*Hint:* type `?mpg` to read the documentation for the dataset). How can you see this information when you run mpg?
Categorical:Manufacturer, model, trans, drv, fl, class 
Continuous:Displ, year, cyl, cty, hwy

```{r}
mpg
```


**3:** Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables? **Note:** You need several code chunks here to answer this question.

A continuous variable cannot be mapped to shape since the difference between a square, circle, triangle is hard to distinguish with a value.

```{r}
ggplot(mpg, aes(x = displ, y = hwy, colour = fl)) +
  geom_point()

ggplot(mpg, aes(x = displ, y = hwy, size = fl)) +
  geom_point()

ggplot(mpg, aes(x = displ, y = hwy, shape = fl)) +
  geom_point()

ggplot(mpg, aes(x = displ, y = hwy, color = cyl)) +
  geom_point()

ggplot(mpg, aes(x = displ, y = hwy, size = cyl)) +
  geom_point()

ggplot(mpg, aes(x = displ, y = hwy, shape = cyl)) +
  geom_point()
```


**4:** What happens if you map the same variable to multiple aesthetics?
This is just a way to reinforce the same information that was proivded and one should be cautious if this occurs.

```{r}
ggplot(mpg, aes(x = displ, y = hwy, color = hwy, size = displ)) +
  geom_point()

```


**5:** What does the stroke aesthetic do? What shapes does it work with? (*Hint:* use `?geom_point`)

Stroke aesthetic defines the width of the border of the shape
It works with shapes :21-25



**6:** What happens if you map an aesthetic to something other than a variable name, like `aes(colour = displ < 5)`?

This can show characteristics for expressions such as true or false.

```{r}
ggplot(mpg, aes(x = displ, y = hwy, color = displ < 5)) +
  geom_point()
```


### Section 3.5: Facets

#### Section 3.5 questions

**1:** What happens if you facet on a continuous variable? **Note:** Write a prediction, then make and run your code chunk.
The facet will display all of the variables

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) + geom_point() + facet_grid(. ~ cty)
```


**2:** What do the empty cells in plot with `facet_grid(drv ~ cyl)` mean? How do they relate to this plot?

These plots which have empty cells do not have values for the combination of drv and cyl.


```{r}
ggplot(data = mpg) + facet_grid(drv ~ cyl)
ggplot(data = mpg) + geom_point(mapping = aes(x = drv, y = cyl))

```


**3:** What plots does the following code make? What does `.` do?
facet plots for drv and cyl, the period ignores the dimensions of which spot it takes

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```


**4:** Take the first faceted plot below. What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?
Advantages, it divides it up by the groups instead of having them all on one chart. Disadvantes are that the different categories cannot be compaired as well if the colors are similar. Larger data set would allow for more accurate grouping and show better overlap.

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```


**5:** Read `?facet_wrap`. What does `nrow` do? What does `ncol` do? What other options control the layout of the individual panels? Why doesn’t `facet_grid()` have nrow and ncol argument?

nrow and ncol: shows the number of rows and columns that the facet will have. the values are already determined in facet grid

**6:** When using `facet_grid()` you should usually put the variable with more unique levels in the columns. Why?

Because there is more room if you use landscape and the graph will be displayed much more nicely
### Section 3.6: Geometric objects



#### Section 3.6.1 Questions

**1:** What geom would you use to draw a line chart? A boxplot? A histogram? An area chart? **Note:** You will have to make some types of these charts in Part 2 of this assignment.


geom_line()
geom_boxplot()
geom_histogram()
geom_area()


**2:** Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.
Color scatterplot with lines for the different asethetics

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + geom_point() + geom_smooth(se = FALSE)
```


**3:** What does `show.legend = FALSE` do? What happens if you remove it?
**Note:** Skip the "Why do you think I used it earlier in the chapter?" question.

Show.legend = False removes the legend and if you remove it then the legend shows.
This wasnt used earlier because the values did not need a legend to describe them.


**4:** What does the `se` argument to `geom_smooth()` do?
Display the confidence interval of the plot around the line

**5:** Will these two graphs look different? Why/why not?
No, they use the same data throughout both plots.
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```


**6:** Recreate the R code necessary to generate the following graphs. **Note:** Enter a separate code chunk for each graph, numbered 1 to 6, starting row-wise at upper left. 

1: upper left.

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) + geom_point() + geom_smooth(se = FALSE)
```


2: upper right.

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) + geom_smooth(mapping = aes(group = drv), se = FALSE) + geom_point()
```


3: middle left.

```{r}
ggplot(mpg, aes(x = displ, y = hwy, colour = drv)) + geom_point() + geom_smooth(se = FALSE)
```


4: middle right.

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) + geom_point(aes(colour = drv)) + geom_smooth(se = FALSE)
```


5: lower left.

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) + geom_point(aes(colour = drv)) + geom_smooth(aes(linetype = drv), se = FALSE)
```


6: lower right. *Hint:* Read the help file for `geom_point`. Use `shape = 21`, `stroke = 2`, and `size = 3`.

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) + geom_point(size = 3, stroke = 2, shape = 21) + geom_point(aes(colour = drv))
```


### Section 3.7: Statistical transformations.

**Skip this section.** We may return to it later.


### Section 3.8 Position adjustments


#### Section 3.8 Questions

**1:** What is the problem with this plot? How could you improve it. *Hint:* Remember the size of the `mpg` data set from earlier in this exercise. Does the number of points seem correct? **Note:** Write the code to improve it. 

This plot is an crowded, it can be improved by a jitter plot

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + geom_point(position = "jitter")
```


**2:** What parameters to `geom_jitter()` control the amount of jittering?

Width and Height control the amount of jittering

**3:** Compare and contracts `geom_jitter()` with `geom_count()`.
Jitter plot doenst overlaps data, while count plays off of the overplotting and uses it to its advantage 

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + geom_jitter()

ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + geom_count()
```



**4:** What is the default position adjustement for `geom_boxplot()`? Create a visualization of the `mpg` dataset that demonstrates it.
The default position is dodge
```{r}
ggplot(data = mpg, aes(x = drv, y = hwy, color = class)) + geom_boxplot()
```



### Section 3.9: Coordinate systems

**Note:** Read about and do only the code chunk associated with `coord_flip()`. `coord_flip()` is a great layer to use when you have categorical data with longish names as it makes the text easier to read.

```{r}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + geom_boxplot() + coord_flip()
```

### Section 3.10: The layered grammar of graphics

Just read this section as a contextual review of everything you just learned.


