---
title: "HW05 Part 1: Complete the sections"
author: "Melvin Cummings"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

###Limpet
```{r}
library(tidyverse)
limpets <- read.csv("~/Bi485/cummings_melvin/Data/limpets.csv")
limpets
```

```{r}
read_csv("~/Bi485/cummings_melvin/Data/limpets.csv", col_types = c("ncc"))
factor(limpets$Site %>% c("PBL", "KNRM", "VBG", "WP", "PF", "DP", "SIO", "CTZ1", "CTZ2", "CNM"), ordered = TRUE)
ggplot(data = limpets, mapping = aes(x = Site, y = Length, color = Protected)) + geom_boxplot() + ylab("Length (mm)") 
```


The box plot shows that the outliers are found within DP and WP. The protected sites have larger quartile ranges then those who were not protected.

##Roseate Terns

```{r}
library(tidyverse)
roseate <- read_tsv("~/Bi485/cummings_melvin/Data/roseate_terns.txt")

T_roseate <- roseate %>% 
  gather("AR", "AL", "AC", "BR", "BL", "BC", "JR", "JL", "JC", key = "Site", value = "People")
T_roseate

filter(T_roseate, is.na(People) == FALSE)

ggplot(data = T_roseate,mapping = aes(x = year,y = People))+geom_line() + facet_wrap(~ Site, nrow = 2)


```

The populations that increased were, AC, Al, AR, BC, and BL, AR had the biggest incresase. The decreasing were Br and JR. While JC and JL stayed relatively the same.

##Darters

```{r}
darter <- read_tsv("~/Bi485/cummings_melvin/data/darter.txt")
darter

start_pos <- c(1, 4, 15, 16, 17, 18, 21, 25, 29, 33, 36, 39, 40)
stop_pos <- c(3, 14, 15, 16, 17, 20, 24, 28, 32, 35, 38, 40, 41)
names <- c("specimen id", "species", "riffle", "age", "sex", "length", "depth", "velocity", "temperature", "majsub", "minsub", "majtype", "mintype")

darter <- read_fwf("darter.txt", fwf_positions(start_pos, stop_pos, col_names = names), skip = 15)

darter_riffle <- darter$riffle <- factor(darter$riffle, levels = c("1", "2"), ordered = FALSE)

darter_majT <- darter$majtype <- factor(darter$majtype, levels = c("s", "fg", "sg", "lg", "c"), ordered = TRUE)

```

```{r}
darter_1 <- darter %>%  filter(!species %in% c("zonale", "tetrazonum")) 
darter_1
```
```{r}
darter_2 <- subset(darter_1, select = -c(mintype, minsub))
darter_2

colnames(darter_2)[colnames(darter_2)=="majsub"] <- "major_substrate"

colnames(darter_2)[colnames(darter_2)=="majtype"] <- "major_type"
darter_2


```

Plot 1
```{r}
ggplot(data = darter_2, mapping = aes(x = length, y = depth, color = species, shape = species)) + 
  geom_point() +
  facet_wrap(~ riffle, nrow = 2)
```
There are fewer outliers in the second riffle

Plot 2
```{r}
ggplot(data = darter_2, mapping = aes(x = species, y = length)) + 
  geom_boxplot() +
  facet_wrap(~ riffle, nrow = 2)
```

Plot 2 shows the greatest number of outliers

Plot 3
```{r}
ggplot(data = darter_2, mapping = aes(x = "majtype", y = length)) +
  geom_boxplot() +
  facet_grid(species ~ riffle) +
  labs(x = "Major Type of Species", y = "Length (cm)")
```

```{r}
ggplot(data = darter_2, mapping = aes(x = "majtype", y = length)) +
  geom_boxplot() +
  facet_grid(riffle ~ species) +
  labs(x = "Major Type of Species", y = "Length (cm)")
```
Changes the chart colums and rows from 2:3 to 3:2

##Blacklip Abalone
Chunk1
```{r}
library(tidyverse)
abalone <- read.csv("~/Bi485/cummings_melvin/Data/abalone.csv")
ggplot(data = abalone) + geom_boxplot(mapping = aes(x = Type, y = Height))
```

Chunk2
```{r}
ggplot(data = abalone) + geom_point(mapping = aes(x = Rings, y = Height
))
```
The large individuals are not related to the old individuals

Chunk3
```{r}
abalone_filter <- filter(abalone, Height < 0.3)

abalone_Type <- abalone <- factor(abalone, levels = c("I", "F", "M"), ordered = TRUE)

ggplot(data = abalone_filter, mapping = aes(x = Rings, y = Height)) + geom_point() + facet_wrap(~ Type, nrow = 1)
```
The pattern that emerge are that females and males are typically larger than immatures. Females seem to be larger in Height than males. Not all immatures are smaller than females or males.

Chunk 4
```{r}
filter(abalone_filter, Height == 0)
abalone_filter[abalone_filter == 0] <- NA
Newabalone_filter <- filter(abalone_filter, is.na(Height) == FALSE)
ggplot(data = Newabalone_filter, mapping = aes(x = Rings, y = Height)) + geom_point() + facet_wrap(~Type, 1)
```

```{r}
ggplot(data = Newabalone_filter, mapping = aes(x = ShellWeight, y = Diameter, shape = Type, color = Rings)) + geom_point()

ggplot(data = Newabalone_filter, mapping = aes(x = Type, y = Diameter, color = Rings)) + geom_jitter()  
```


