---
title: "HW 10"
author: "Melvin Cummings"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
library(raster)
library(tidyverse)
library(ggthemes)
library(patchwork)
library(maptools)
library(ggmap)
library(marmap)
```

```{r}
usa <- map_data(map= "state")
usa
```


```{r}
us_missouri <- map_data(map = "state",
                        region = "missouri")

us_missouri
```

```{r}
us_other <- subset(x = usa,
subset = region %in% c("alabama",
"florida",
"mississippi",
"north carolina"))

us_other
```


```{r}
ggplot() + 
  geom_polygon(data = us_other, 
               aes(x = long, 
                   y = lat, 
                   group = group), 
               fill = "gray")
```

```{r}
ggplot() +
geom_polygon(data = us_other,
aes(x = long,
y = lat,
group = group),
fill = "gray") +
geom_polygon(data = us_missouri,
aes(x = long, y = lat,
group = group),
fill = "black")
```

```{r}
ggplot() +
geom_polygon(data = us_other,
aes(x = long,
y = lat,
group = group),
fill = "gray") +
geom_polygon(data = us_missouri,
aes(x = long, y = lat,
group = group),
fill = "black") +
geom_polygon(data = usa,
aes(x=long,
y = lat,
group = group),
fill = NA,
color = "black")
```

```{r}
us_map <- ggplot() +
geom_polygon(data = us_other,
aes(x = long,
y = lat,
group = group),
fill = "gray") +
geom_polygon(data = us_missouri,
aes(x = long, y = lat,
group = group),
fill = "black") +
geom_polygon(data = usa,
aes(x=long,
y = lat,
group = group),
fill = NA,
color = "black") +
theme_map() +
coord_fixed(1.3)

us_map
```

```{r}
missouri <- map_data("county", "missouri")
mo_scott <- subset(missouri, subregion %in% c("scott"))
mo_map <- ggplot() +
geom_polygon(data = missouri, aes(x=long, y = lat, group = group), fill = NA, color = "black") +
geom_polygon(data = mo_scott, aes(x = long, y = lat, group = group), fill = "black") +
theme_map() +
coord_fixed(1.3)

mo_map
```

```{r}
us_map + mo_map + plot_layout(ncol = 2, widths = c(1.5,1))

```

```{r}
us_map_melvin <- ggplot() + 
  geom_polygon(data = us_missouri, 
               aes(x = long, y = lat, 
                   group = group), 
               fill = "black") +
  geom_polygon(data = usa, 
               aes(x=long, 
                   y = lat, 
                   group = group), 
               fill = NA, 
               color = "black") + 
  theme_map() +
  coord_fixed(1.3)
us_map_melvin
```

```{r}
missouri <- map_data("county", "missouri")
mo_birth <- subset(missouri, subregion %in% c("pemiscot"))
mo_names <- subset(missouri, subregion %in% c("madison", "cole"))

missouri_melvin <- ggplot()+
  geom_polygon(data = mo_birth,
               aes(x = long,
                   y = lat,
                   group = group),
               fill = "black")+
  geom_polygon(data = mo_names,
               aes(x = long,
                   y = lat,
                   group = group),
               fill = "gray")+
  geom_polygon(data = missouri,
               aes(x = long,
                   y = lat,
                   group = group),
               fill = NA,
               color = "black")+
  theme_map()+
  coord_fixed(1.3)

missouri_melvin
```


```{r}
us_map_melvin + missouri_melvin + plot_layout(ncol = 2, widths = c(1.5,1))
```

#Part 2

```{r}
library(raster)
library(tidyverse)
library(ggthemes)
library(ggmap)
```

```{r}
min_long <- -126
max_long <- -104
min_lat  <-   31
max_lat  <-   50

set.seed(9618973)
```

```{r}
file_path <- "~/Bi485/cummings_melvin/data"

bigfoot <-
  read_csv(file.path(file_path, "bigfoot.csv"))
bigfoot
```

```{r}
bigfoot <- filter(bigfoot,
                  long >= min_long & long <= max_long,
                  lat >= min_lat & lat <= max_lat) %>% 
  sample_n(300) %>% 
  mutate(name = "Gigantopithecus")
bigfoot

```
```{r}
file_path2 <- "~/Bi485/cummings_melvin/data"

bears <-
  read_csv(file.path(file_path, "bears.csv"))
bears
```



```{r}
both_species <- bind_rows(bigfoot, bears)
head(both_species)
```

```{r}
tail(both_species)
```

```{r}
base = get_stamenmap(bbox = c(min_long,
min_lat,
max_long,
max_lat),
zoom = 4,
maptype = "terrain-background")
ggmap(base)
```

```{r}
base <- ggmap(base)
base + geom_point(data = bigfoot,
aes(x = long,
y = lat))
```

```{r}
base + geom_point(data = bears,
aes(x = long,
y = lat))
```

```{r}
base + geom_point(data = both_species,
aes(x = long,
y = lat,
shape = name,
fill = name),
size = 2) +
scale_shape_manual(values = c(22:23)) +
labs(x = "Longitude",
y = "Latitude",
fill = "Species",
shape = "Species")

```

```{r}
min_long <- -126
max_long <- -104
min_lat  <-   31
max_lat  <-   50
set.seed(1790569)
```

```{r}

```{r}
file_path <- "~/Bi485/cummings_melvin/data"

bigfoot <-
  read_csv(file.path(file_path, "bigfoot.csv"))
bigfoot
```

```{r}
bigfoot <- filter(bigfoot,
                  long >= min_long & long <= max_long,
                  lat >= min_lat & lat <= max_lat) %>% 
  sample_n(300) %>% 
  mutate(name = "Gigantopithecus")
bigfoot

```
```{r}
file_path2 <- "~/Bi485/cummings_melvin/data"

bears <-
  read_csv(file.path(file_path, "bears.csv"))
bears
```



```{r}
both_species <- bind_rows(bigfoot, bears)
head(both_species)
```

```{r}
tail(both_species)
```

```{r}
base = get_stamenmap(bbox = c(min_long,
min_lat,
max_long,
max_lat),
zoom = 4,
maptype = "terrain-background")
ggmap(base)
```

```{r}
base <- ggmap(base)
base + geom_point(data = bigfoot,
aes(x = long,
y = lat))
```

```{r}
base + geom_point(data = bears,
aes(x = long,
y = lat))
```

```{r}
base + geom_point(data = both_species,
aes(x = long,
y = lat,
shape = name,
fill = name),
size = 2) +
scale_shape_manual(values = c(22:23)) +
labs(x = "Longitude",
y = "Latitude",
fill = "Species",
shape = "Species")

```


```{r}
min_long <- -170
max_long <- 164
min_lat <- -42
max_lat <- -8

earthquakes <- quakes %>%
mutate(depth = depth * 3280.84)
```

```{r}
fiji <- getNOAA.bathy(lon1 = min_long,
lon2 = max_long,
lat1 = min_lat,
lat2 = max_lat,
antimeridian = TRUE,
keep = TRUE)

```

```{r}
autoplot.bathy(fiji,
geom = c("raster", "contour"),
size = 0.1,
na.rm = TRUE)
```

```{r}
autoplot(fiji,
geom = c("raster","contour"),
size = 0.1,
na.rm = TRUE) +
scale_fill_gradient2(low = "dodgerblue4",
mid = "gainsboro",
high = "darkgreen",
name = "Depth (ft)") +
labs(x = "Longitude",
y = "Latitude") +
theme(axis.title.y = element_text(angle = 0,
vjust = 0.5))
```

```{r}
base_map <- autoplot(fiji,
geom = c("raster","contour"),
size = 0.1,
na.rm = TRUE) +
scale_fill_gradient2(low = "dodgerblue4",
mid = "gainsboro",
high = "darkgreen",
name = "Depth (ft)") +
labs(x = "Longitude",
y = "Latitude") +
theme(axis.title.y = element_text(angle = 0,
vjust = 0.5))
```

```{r}
base_map +
geom_point(data = earthquakes,
aes(x = long,
y = lat))
```

```{r}
base_map +
geom_point(data = earthquakes,
aes(x = long,
y = lat,
size = mag),
alpha = 0.4) +
scale_size_continuous(range = c(0.5,3), name = "Magnitude")
```

```{r}
min_long <- -90
max_long <- -58
min_lat <- 8
max_lat <- 28
```

```{r}
file_path3 <-"~/Bi485/cummings_melvin/data"
blennies <- read.csv(file.path(file_path3, "blennies.csv"))
blennies
```

```{r}
carib_sea <- getNOAA.bathy(lon1 = min_long,
                           lon2 = max_long,
                           lat1 = min_lat,
                           lat2 = max_lat,
                           antimeridian = FALSE,
                           keep = TRUE)
```

```{r}
base_map <- autoplot(carib_sea,
         geom = c("raster", "contour"),
         size = 0.1,
         na.rm = TRUE) +
  scale_fill_etopo(guide = FALSE)+
  labs(x = "Longitude",
       y = "Latitude")+
  theme(axis.title.y = element_text(angle = 0,
                                    vjust = 0.5))
base_map
```

```{r}
base_map + geom_point(data = blennies,
                      aes(x = long,
                          y = lat,
                          shape = name),
                      fill = "white",
                      size =2.5)+
  scale_shape_manual(values = c(21:23))+
  labs(shape = "Species")+
  theme(legend.position = "bottom")+
  theme(legend.text = element_text(face = "italic"))

```


