---
title: "HW 06 Part 1: Complete the sections"
author: "Melvin Cummings"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

- Change "your name" in the YAML header above to your name.

- As usual, enter the examples in code chunks and run them, unless told otherwise.

## Chapter 5: Data transformation

Read [R4ds Chapter 5: Data transformation](https://r4ds.had.co.nz/transform.html), sections 5-7.

You do not answer any of the questions this week. Just read and run all of the examples.

### Section 5.5: Add new variables with `mutate()`

Type and run all of the examples.

```{r}
library(nycflights13)
library(tidyverse)
flights_sml <- select(flights,
                      year:day,
                      ends_with("delay"),
                      distance,
                      air_time
                      )
mutate(flights_sml,
       gain = dep_delay - arr_delay,
       speed = distance / air_time * 60
       )
```

Mutate:Adds a new column

```{r}
transmute(flights,
          gain = dep_delay - arr_delay,
          hours = air_time / 60,
          gain_per_hour = gain / hours
          )
```

Transmute: keeps new variables

```{r}
transmute(flights,
          dep_time,
          hour = dep_time %/% 100,
          minutes = dep_time %% 100
          )
```

```{r}
(x <- 1:10)

lag(x)

lead(x)
```

```{r}
x

cumsum(x)

cummean(x)
```

```{r}
y <- c(1, 2, 2, NA, 3, 4)

min_rank(y)

min_rank(desc(y))

row_number(y)

dense_rank(y)

percent_rank(y)

cume_dist(y)
```



Pay careful attention to the useful creation functions in section 5.5.1. 

### Section 5.6: Grouped summaries with `summarise()`

```{r}
summarise(flights,
          delay = mean(dep_delay, na.rm = TRUE))

by_day <- group_by(flights,
                   year,
                   month,
                   day)

summarise(by_day, delay = mean(dep_delay, na.rm = TRUE))
```

```{r}
by_dest <- group_by(flights, dest)
delay <- summarise(by_dest,
                   count = n(),
                   dist = mean(distance, na.rm = TRUE),
                   delay = mean(arr_delay, na.rm = TRUE)
                   )
delay <- filter(delay, count > 20, dest != "HNL")

ggplot(data = delay, mapping = aes(x = dist, y = delay)) +
  geom_point(aes(size = count), alpha = 1/3) +
  geom_smooth(se = FALSE)
```

```{r}
delays <- flights %>% 
  group_by(dest) %>% 
  summarise(
    count = n(),
    dist = mean(distance, na.rm = TRUE),
    delay = mean(arr_delay, na.rm = TRUE)
  ) %>% 
filter(count > 20, dest != "HNL")
```

```{r}
flights %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay))
```

na.rm: removes the NA values

```{r}
flights %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay, na.rm = TRUE))
```
```{r}
not_cancelled <- flights %>%
  filter(!is.na(dep_delay), !is.na(arr_delay))

not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay))
```



Read and run all of the examples. Pay careful attention to the useful summary functions. Unless you really like baseball, skip the baseball part of 5.6.3.

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(
    avg_delay1 = mean(arr_delay),
    avg_delay2 = mean(arr_delay[arr_delay > 0])  
  )
```

```{r}
not_cancelled %>% 
  group_by(dest) %>% 
  summarise(distance_sd = sd(distance)) %>% 
  arrange(desc(distance_sd))
```

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(
    first = min(dep_time),
    last  = max(dep_time)
  )
```

```{r}
not_cancelled %>% 
  group_by(year,month,day) %>% 
  mutate(r=min_rank(desc(dep_time))) %>% 
  filter(r %in% range(r))
```

```{r}
not_cancelled %>% 
  group_by(dest) %>% 
  summarise(carrier = n_distinct(carrier)) %>% 
  arrange(desc(carrier))
```

```{r}
not_cancelled %>% 
  count(dest)
```

```{r}
not_cancelled %>% 
  count(tailnum, wt = distance)
```

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(n_early = sum(dep_time < 500))

not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(hour_perc = mean(arr_delay > 60))

```

```{r}
daily <- group_by(flights, year, month, day)
(per_day   <- summarise(daily, flights = n()))

(per_month <- summarise(per_day, flights = sum(flights)))

(per_year  <- summarise(per_month, flights = sum(flights)))
```

```{r}
daily %>% 
  ungroup() %>% 
  summarise(flights = n())

```



### Section 5.7: Grouped mutates (and filters)

Read and run all of the examples.

```{r}
flights_sml %>% 
  group_by(year, month, day) %>%
  filter(rank(desc(arr_delay)) < 10)
```

```{r}
popular_dests <- flights %>% 
  group_by(dest) %>% 
  filter(n() > 365)
popular_dests
```

```{r}
popular_dests %>% 
  filter(arr_delay > 0) %>% 
  mutate(prop_delay = arr_delay / sum(arr_delay)) %>% 
  select(year:day, dest, arr_delay, prop_delay)
```
